# week 3:

loop functions

lapply sapply
tapply mapply 
apply: let's set margins


gl

debbuger
* message: informative 
* warnings: unexpected behavior, display once execution has finished
* error: fatal breaks the program
* condition: generic



out printing to the console

useful tools


traceback
debug 
browser
trace 
recover

blunt techniques insert cat/ print staments insid the function

```{r eval=FALSE}
mean(x)
traceback()

debug(lm)
lm(x-y)

undebug(lm)

optins(error=recover)
optins(error=browser)

optins(error=NULL)

```




 
```{r eval=FALSE}
tidyverse_update()
```


<<- is an operator used to assign a object to a different enviroment




## R objects
## Create package documentation 
Run ```devtools::document()``` (or press Ctrl/Cmd + Shift + D in RStudio) to convert roxygen comments to .Rd files. 



## SWIRL practices 

### lapply sapply
loop functions or *apply functions

__split apply combine__ strategy lay out by Hadley Wickham in his  paper titled 'The Split-Apply-Combine Strategy for Data Analysis'.



Rows observation and columns variables


especiffy its own anonimus function 
lapply(unique_vals, function(elem) elem[2])

back in week to we read some data from a file. Imagine we want to analize this information we want to know what type is. first we try simply to find the class of students which ain't very us 

Now we want to iterate operation over it
lapply returns a list type as a result. sapply tries to simplify whenever is possible. iF every result is of length 1 it returns a vector if they are all bigger than one and with the same length it returns a matrix. 
```{r}
students <- read.table("../data/student_info.csv",sep=";",header=TRUE,nrows=5)
class(students)

cls<- lapply(students,class)

```



### vapply and tapply
vapply allows to specify the format of the result and throughs an error if not matched.




more in depth of the source function
source from web github
```{r}

print(" I am filed called from a URL")
source("https://raw.githubusercontent.com/jsduenass/datasciencecoursera/master/code/hello_I_am_from_the_web.R")

```


One can call an R script and all its content using source. R can also interface with different files compress files using ... and web pages using the url function.
```{r eval=FALSE}
#remove(list=ls())
#source("code/simple_script.R", encoding = 'UTF-8')
#eval(parse("simple_script.R", encoding="UTF-8"))
#roll()
#dput(rank)
#dump(c(rank,rolls))
```


```{r eval=TRUE}
content <- file("../data/student_info.csv","r")
some_lines <-readLines(content,5)
print(some_lines)
```
 


r measure run time and optimization

