# week 4: Simulation and modeling

proliferer in R a tool that allows you to check the performance of your functions and would help you optimize 

The ```str()``` function gives a quick examination of R objects functions and data perrty usefull to have a general idea of what an specific expression is. 

```{r eval=FALSE}
str(str)
library(datasets)
head(airquality)
str(airquality)
s <- split(airquality,airquality$Month)
str(s) 
```


## simulation 
because R is statistics orientated it has many functions used for creating distributions.  Each type of distribution like normal, Poisson, etc has asociated the following prefixes that specify different outcomes.

* __d:__ density
* __r:__random number generation
* __p:__ cumulative
* __q:__ quantile
 
 The following are some examples.
 
* __rnorm:__ normal distribution given a mean and a standar deviation 

* __dnorm:__ Probability density

* __pnorm:__ Cumulative distribution fro a normal distribution 

*__rpois:__ generates a random Possion distribution given a defined rate

```{r}

with(airquality,tapply(Temp,Month,mean))
Temperature<- airquality$Temp

model <-rnorm(153,mean(Temperature),sd(Temperature))
 

plot(density(Temperature))

plot(density(model))

```

R produces random number that are not entierly random, instead they are made through a pseudo random generator, an algorithm that will map the output to a list of values with a determined probabilistic distribution. A seed will determine the behavior of the algorithm, so when a seed has been set it will always map to the same numbers. It is important to be able to reproduce the exact same results, so the ```set.seed()``` function is used to ensure that the random values generated stay the same.



```{r}
set.seed(3)
```


